\bf Visualising How Non-linear Dimension Reduction Warps Your Data

### 2024-01-10 paper.qmd
Added figure titles

p1 <-  ggplot(data = hex_full_count_df_loop1, aes(x = x, y = y)) +
      geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
            geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
      scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
        theme_void() +
        theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
              legend.title = element_text(size=8), #change legend title font size
              legend.text = element_text(size=6)) +
        guides(fill = guide_colourbar(title = "Standardized count")) +
        annotate(geom = 'text', label = "a", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)

@@

p2 <-  ggplot(data = hex_full_count_df_loop2, aes(x = x, y = y)) +
  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
    geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
        theme_void() +
    theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
          axis.title.x = element_blank(), axis.title.y = element_blank(),
          axis.text.x = element_blank(), axis.ticks.x = element_blank(),
          axis.text.y = element_blank(), axis.ticks.y = element_blank(),
          panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
          legend.title = element_text(size=8), #change legend title font size
          legend.text = element_text(size=6)) +
    guides(fill = guide_colourbar(title = "Standardized count")) +
    annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)

@@

```{r}
#| warning: false
#| echo: false
#| label: fig-wkhighD
#| fig-cap: Visualization.

a1 + a2 +
  plot_annotation(tag_levels = 'a') +
  plot_layout(guides='collect', ncol = 3) &
  theme(legend.position='bottom', plot.tag = element_text(size = 8))
```

@@

## To draw with AIC
aic_plot <- ggplot(MSE_df |> dplyr::filter(data_type == "training"), aes(x = number_of_bins,
                                                                                 y = total_error,
                                                                                 color = method
)) +
  geom_point() +
  geom_line() +
  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
  theme_light() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 7)) +
  scale_color_discrete_qualitative() +
  ylab("AIC") +
  xlab("Total number of bins")
## Effective number of bins along x-axis

mse_plot <- ggplot(MSE_df, aes(x = number_of_bins,
                                       y = total_mse,
                                       color = method
)) +
  geom_point() +
  geom_line() +
  theme_light() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 7)) +
  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
  scale_color_discrete_qualitative() +
  ylab("MSE") +
  xlab("Total number of bins")

@@

```{r}
#| echo: false
#| fig-cap: The .
#| label: fig-diagnosticplt
#| out-width: 100%

@@

mse_plot_param_training <- ggplot(MSE_df_umap |> dplyr::filter(data_type == "training"), aes(x = number_of_bins,
                               y = total_mse,
                               color = param
)) +
  geom_point() +
  geom_line() +
  theme_light() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 7)) +
  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
  scale_color_discrete_qualitative() +
  ylab("MSE") +
  xlab("Total number of bins")

mse_plot_param_test <- ggplot(MSE_df_umap |> dplyr::filter(data_type == "test"), aes(x = number_of_bins,
                                                                                             y = total_mse,
                                                                                             color = param
)) +
  geom_point() +
  geom_line() +
  theme_light() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 7)) +
  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
  scale_color_discrete_qualitative() +
  ylab("MSE") +
  xlab("Total number of bins")

@@

```{r}
#| echo: false
#| fig-cap: The .
#| label: fig-diagnosticpltDiffParam
#| out-width: 100%

@@

<!--In this analysis, we're looking at how well our model is doing in two aspects. First, we're checking the distances in our lower-dimensional space (see @fig-diag (a)). We want to see if the distances between different types of edges are similar. If they are, it means our model is doing a good job of keeping the relationships consistent. If there are big differences, it could mean our model is struggling with certain types of edges.

Next, we're examining how our model handles the relationship between high-dimensional and low-dimensional distances for each type of edge (see @fig-diag (b)). If our data points nicely line up around the expected line, it tells us that our model is preserving the original structure well. If they're scattered, it could mean our model is having trouble with specific data points. This two-part analysis helps us really understand how our model fits the data.

sorted_distance_df <- distance[order(distance$distance),] ## Sort the distances

sorted_distance_df <- cbind(sorted_distance_df, rbind(NA, apply(sorted_distance_df, 2, diff))) ## Calculate differences between unique distance
sorted_distance_df <- sorted_distance_df[, c(3,6)]
sorted_distance_df[is.na(sorted_distance_df)] <- 0.0000 ## To replace missing values with zero
names(sorted_distance_df)[2] <- "difference"

sorted_distance_df <- sorted_distance_df %>% 
  mutate(Edge_type = ifelse(distance >= benchmark, "Long edge", "Short edge"))

diagnostic_plot1 <- ggplot(sorted_distance_df, aes(x = distance, y = difference, color = Edge_type)) +
  geom_point(alpha = 0.5) +
  geom_rug(col="#e31a1c",alpha=0.1, linewidth=1.5) +
  #ggtitle("Visualization of difference between \n 2D distances by edge type.") +
  scale_color_manual(values=c("#33a02c", "#000000")) + 
  theme(legend.position = "bottom", legend.title=element_blank(), plot.title=element_text(size=8), axis.text = element_text(size = 5),
          axis.title = element_text(size = 7),
        legend.text = element_text(size=7)) + 
  xlab(expression(d^{(2)})) 
```

distance <- distance %>% 
  rename(distance_2D = distance)

#num_pca <- 4

high_d <- df_bin[, 1:7]

data_matrix <- data.matrix(high_d, rownames.force = NA)

# calculate Euclidean distance between each row in matrix
d <- dist(data_matrix)

df <- reshape2::melt(as.matrix(d), varnames = c("Start_point", "End_point"))

df <- df[df$Start_point > df$End_point,]

rownames(df) <- 1:nrow(df)

df <- df %>% 
  rename(distance_high_D = value)

high_low_dist <- left_join(distance, df, by = c("from" = "End_point", "to" = "Start_point"))
high_low_dist <- high_low_dist %>% distinct()
high_low_dist <- high_low_dist[complete.cases(high_low_dist),]

lm_fit <- lm(distance_high_D ~ distance_2D, data = high_low_dist)

high_low_dist <- high_low_dist %>% 
  mutate(Edge_type = ifelse(distance_2D >= benchmark, "Long edge", "Short edge"))

diagnostic_plot2 <- ggplot(high_low_dist, aes(x = distance_2D, y = distance_high_D, color = Edge_type)) +
  geom_point(alpha = 0.5) +
  geom_line(data = broom::augment(lm_fit),
            aes(x = distance_2D, y = .fitted),
            color="blue",
            linewidth=1) +
  geom_rug(col="#e31a1c",alpha=0.1, linewidth=1.5) +
  #ggtitle("Visualization of 2D and high-D distances \n by edge type.") +
  scale_color_manual(values=c("#33a02c","#000000")) + 
  theme(legend.position = "bottom", legend.title=element_blank(), plot.title=element_text(size=8), axis.text = element_text(size = 5),
          axis.title = element_text(size = 7),
        legend.text = element_text(size=7)) + 
  xlab(expression(d^{(2)})) + 
  ylab(expression(d^{(p)}))
```

```{r}
#| echo: false
#| fig-cap: The two diagnostic plot assess the effectiveness of the high-D representation algorithm  performed for the **s_curve_noise_umap** data set$\colon$ (a) visualizing the differences between low-D distances and (b) visualizing the relationship between high-D ($d^{(p)}$) and low-D distances ($d^{(2)}$). 
#| label: fig-diag
#| out-width: 100%

diagnostic_plot1 + diagnostic_plot2 +
  plot_annotation(tag_levels = 'a') +
  plot_layout(guides='collect') &
  theme(legend.position='bottom', plot.tag = element_text(size = 8))
```

### Software

Our algorithm for high-dimensional (high-D) representation is available as an open-source R package named quollr. This package provides analysts with the capability to construct a low-dimensional manifold and seamlessly visualize it in the high-D space alongside the original data.-->

@@

## To draw with AIC
aic_gau_plot <- ggplot(MSE_df_gau |> dplyr::filter(data_type == "training"), aes(x = number_of_bins,
                                                             y = total_error,
                                                             color = method
)) +
  geom_point() +
  geom_line() +
  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
  theme_light() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 7)) +
  scale_color_discrete_qualitative() +
  ylab("AIC") +
  xlab("Total number of bins")
## Effective number of bins along x-axis

mse_gau_plot <- ggplot(MSE_df_gau, aes(x = number_of_bins,
                   y = total_mse,
                   color = method
)) +
  geom_point() +
  geom_line() +
  theme_light() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 7)) +
  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
  scale_color_discrete_qualitative() +
  ylab("MSE") +
  xlab("Total number of bins")

@@

@@

::: {#fig-pbmc_sc layout-ncol="3"}
![](figures/pbmc_a.png){#fig-pbmc_sc1}

@@

Screen shots of the **langevitour** with the low-dimensional view applied to pbmc data set, a video of the tour animation is available at (\url{https://drive.google.com/file/d/1F7Xb2EIpfUruCxYihPw2rKro4FB5Y2uL/view}). Combination of the **langevitour** and low-dimensional representation of the pbmc data provides the effectiveness of our approach.
:::

@@

>>>>

p1 <-  ggplot(data = hex_full_count_df_loop1, aes(x = x, y = y)) +
      geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
            geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
      scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
  xlim(-5, 7) +
  ylim(-10, 10) +
        theme_void() +
        theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
              axis.title.x = element_blank(), axis.title.y = element_blank(),
              axis.text.x = element_blank(), axis.ticks.x = element_blank(),
              axis.text.y = element_blank(), axis.ticks.y = element_blank(),
              panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
              legend.title = element_text(size=8), #change legend title font size
              legend.text = element_text(size=6)) +
        guides(fill = guide_colourbar(title = "Standardized count")) +
        annotate(geom = 'text', label = "a", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)

@@

p2 <-  ggplot(data = hex_full_count_df_loop2, aes(x = x, y = y)) +
  geom_polygon(color = "black", aes(group = polygon_id, fill = std_counts)) +
    geom_point(data = UMAP_data, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "C") +
    xlim(-5, 8) +
  ylim(-10, 10) +
        theme_void() +
    theme(legend.position="none", legend.direction="horizontal", plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
          axis.title.x = element_blank(), axis.title.y = element_blank(),
          axis.text.x = element_blank(), axis.ticks.x = element_blank(),
          axis.text.y = element_blank(), axis.ticks.y = element_blank(),
          panel.grid.major = element_blank(), panel.grid.minor = element_blank(), #change legend key width
          legend.title = element_text(size=8), #change legend title font size
          legend.text = element_text(size=6)) +
    guides(fill = guide_colourbar(title = "Standardized count")) +
    annotate(geom = 'text', label = "b", x = -Inf, y = Inf, hjust = -0.3, vjust = 1, size = 3)

@@

```{r}
#| warning: false
#| echo: false
#| label: fig-wkhighD
#| fig-cap: How the 2D model lift into high dimensions? (a) visualize the points and the hexagonal bin centroids related 31th and 239th hexagons, (b) visualization of the edge connected the 31th and 239th hexagons (colored in red) in the triangular mesh. A video of tour animation is available at (need to add the link of lg_high).

a2 + a1 +
  plot_annotation(tag_levels = 'a') +
  plot_layout(ncol = 3) &
  theme(legend.position='bottom', plot.tag = element_text(size = 8))
```

@@

## To draw with AIC
aic_plot <- ggplot(MSE_df |> dplyr::filter(data_type == "training"), aes(x = number_of_bins,
                                                                                 y = total_error,
                                                                                 color = method
)) +
  geom_point() +
  geom_line() +
  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
  theme_light() +
  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 7)) +
  scale_color_discrete_qualitative() +
  ylab("AIC") +
  xlab("Total number of bins")
## Effective number of bins along x-axis

mse_plot <- ggplot(MSE_df, aes(x = number_of_bins,
                                       y = total_mse,
                                       color = method
)) +
  geom_point() +
  geom_line() +
  theme_light() +
  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 7)) +
  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
  scale_color_discrete_qualitative() +
  ylab("MSE") +
  xlab("Total number of bins")

@@

```{r}
#| echo: false
#| fig-cap: Goodness of fit statistics from different NLDR techniques applied to training S-curve dataset. What is the best NLDR technique to represent the original data in 2D?
#| label: fig-diagnosticpltScurve
#| out-width: 100%

@@

mse_plot_param_training <- ggplot(MSE_df_umap |> dplyr::filter(data_type == "training"), aes(x = number_of_bins,
                               y = total_mse,
                               color = param
)) +
  geom_point() +
  geom_line() +
  theme_light() +
  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 7)) +
  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
  scale_color_discrete_qualitative() +
  ylab("MSE") +
  xlab("Total number of bins")

mse_plot_param_test <- ggplot(MSE_df_umap |> dplyr::filter(data_type == "test"), aes(x = number_of_bins,
                                                                                             y = total_mse,
                                                                                             color = param
)) +
  geom_point() +
  geom_line() +
  theme_light() +
  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 7)) +
  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
  scale_color_discrete_qualitative() +
  ylab("MSE") +
  xlab("Total number of bins")

@@

```{r}
#| echo: false
#| fig-cap: Goodness of fit statistics from different n_neighbors parameter of UMAP applied to training S-curve dataset. What is the best parameter choice in UMAP to represent the original data in 2D?
#| label: fig-diagnosticpltDiffParam
#| out-width: 100%

@@

@@

## To draw with AIC
aic_gau_plot <- ggplot(MSE_df_gau |> dplyr::filter(data_type == "training"), aes(x = number_of_bins,
                                                             y = total_error,
                                                             color = method
)) +
  geom_point() +
  geom_line() +
  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
  theme_light() +
  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 7)) +
  scale_color_discrete_qualitative() +
  ylab("AIC") +
  xlab("Total number of bins")
## Effective number of bins along x-axis

mse_gau_plot <- ggplot(MSE_df_gau, aes(x = number_of_bins,
                   y = total_mse,
                   color = method
)) +
  geom_point() +
  geom_line() +
  theme_light() +
  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 7)) +
  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
  scale_color_discrete_qualitative() +
  ylab("MSE") +
  xlab("Total number of bins")

@@

```{r}
#| echo: false
#| fig-cap: Goodness of fit statistics from different NLDR techniques applied to training five spherical Gaussian cluster dataset. What is the best NLDR technique to represent the original data in 2D?
#| label: fig-diagnosticpltGau
#| out-width: 100%

@@

<!--
::: {#fig-pbmc_sc layout-ncol="3"}
![](figures/pbmc_a.png){#fig-pbmc_sc1}

@@

Screen shots of the **langevitour** with the low-dimensional view applied to pbmc data set, a video of the tour animation is available at (\url{https://drive.google.com/file/d/1F7Xb2EIpfUruCxYihPw2rKro4FB5Y2uL/view}). Combination of the **langevitour** and low-dimensional representation of the pbmc data provides the effectiveness of our approach.
:::-->

MSE_df_pbmc <- read_rds("data/pbmc/summary_pbmc.rds")

## To draw with AIC
aic_plot_pbmc <- ggplot(MSE_df_pbmc |> dplyr::filter(data_type == "training"), aes(x = number_of_bins,
                                                                                 y = total_error,
                                                                                 color = method
)) +
  geom_point() +
  geom_line() +
  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
  theme_light() +
  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 7)) +
  scale_color_discrete_qualitative() +
  ylab("AIC") +
  xlab("Total number of bins")
## Effective number of bins along x-axis

mse_plot_pbmc <- ggplot(MSE_df_pbmc, aes(x = number_of_bins,
                                       y = total_mse,
                                       color = method
)) +
  geom_point() +
  geom_line() +
  theme_light() +
  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 7)) +
  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
  scale_color_discrete_qualitative() +
  ylab("MSE") +
  xlab("Total number of bins")

```{r}
#| echo: false
#| fig-cap: Goodness of fit statistics from different NLDR techniques applied to training PBMC dataset. What is the best NLDR technique to represent the original data in 2D?
#| label: fig-diagnosticpltPBMC
#| out-width: 100%

aic_plot_pbmc + mse_plot_pbmc +
  plot_annotation(tag_levels = 'a') +
  plot_layout(guides='collect', ncol = 2) &
  theme(legend.position='bottom', plot.tag = element_text(size = 8))
```

@@

MSE_df_zei <- read_rds("data/zeisel/summary_zei.rds")

## To draw with AIC
aic_plot_zei <- ggplot(MSE_df_zei |> dplyr::filter(data_type == "training"), aes(x = number_of_bins,
                                                                                 y = total_error,
                                                                                 color = method
)) +
  geom_point() +
  geom_line() +
  #geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
  #annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=-5000, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
  theme_light() +
  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 7)) +
  scale_color_discrete_qualitative() +
  ylab("AIC") +
  xlab("Total number of bins")
## Effective number of bins along x-axis

mse_plot_zei <- ggplot(MSE_df_zei, aes(x = number_of_bins,
                                       y = total_mse,
                                       color = method
)) +
  geom_point() +
  geom_line() +
  theme_light() +
  theme(legend.title = element_blank(), plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5),
        axis.title = element_text(size = 7),
        axis.text = element_text(size = 7)) +
  # geom_vline(xintercept = NROW(full_grid_with_hexbin_id)) +
  # annotate("text", x= (NROW(full_grid_with_hexbin_id) - 10), y=0.25, label=paste0("effective number of bins = ", as.character(NROW(full_grid_with_hexbin_id))), angle=90) +
  scale_color_discrete_qualitative() +
  ylab("MSE") +
  xlab("Total number of bins")

```{r}
#| echo: false
#| fig-cap: Goodness of fit statistics from different NLDR techniques applied to training Zeisel mouse brain dataset. What is the best NLDR technique to represent the original data in 2D?
#| label: fig-diagnosticpltZEI
#| out-width: 100%

aic_plot_zei + mse_plot_zei +
  plot_annotation(tag_levels = 'a') +
  plot_layout(guides='collect', ncol = 2) &
  theme(legend.position='bottom', plot.tag = element_text(size = 8))
```

### 2024-01-10 paper.tex
Added figure titles

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, enhanced, breakable, interior hidden, frame hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners]}{\end{tcolorbox}}\fi

@@

\caption{\label{fig-wkhighD}Visualization.}

@@

{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-diagnosticplt-1.pdf}

@@

\caption{\label{fig-diagnosticplt}The .}

@@

\caption{\label{fig-diagnosticpltDiffParam}The .}

@@

\includegraphics{paper_files/figure-pdf/unnamed-chunk-43-1.pdf}

@@

Next, we employed the UMAP technique with default parameter settings. As
illustrated in Figure~\ref{fig-pbmc}, the cell types B and Platelet are
well-separated in the UMAP layout. Moreover, CD14+ Mono, FCGR3A+ Mono,
and DC form a distinct cluster, while Naive CD4 T, NK, Memory CD4 T, and
CD8 T are grouped together in another cluster. The values utilized to
construct the smooth low-dimensional manifold are presented in
\textbf{?@tbl-table02}. The linked video, demonstrating the tour with
the model, showcases the generation of a smooth surface for this
application, enabling a comprehensive exploration of the data's
structure and relationships (see Figure~\ref{fig-pbmc_sc}).

@@

\begin{minipage}[t]{0.33\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics{figures/pbmc_a.png}

\subcaption{\label{fig-pbmc_sc1}}
\end{minipage}%
%
\begin{minipage}[t]{0.33\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics{figures/pbmc_b.png}

\subcaption{\label{fig-pbmc_sc2}}
\end{minipage}%
%
\begin{minipage}[t]{0.33\linewidth}

{\centering

\raisebox{-\height}{

\includegraphics{figures/pbmc_c.png}

\subcaption{\label{fig-pbmc_sc3}}
\end{minipage}%

\caption{\label{fig-pbmc_sc}Screen shots of the \textbf{langevitour}
with the low-dimensional view applied to pbmc data set, a video of the
tour animation is available at
(\url{https://drive.google.com/file/d/1F7Xb2EIpfUruCxYihPw2rKro4FB5Y2uL/view}).
Combination of the \textbf{langevitour} and low-dimensional
representation of the pbmc data provides the effectiveness of our
approach.}

@@

\caption{\label{fig-diagnosticpltPBMC}The .}

@@

\caption{\label{fig-diagnosticpltZEI}The .}

>>>>

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, interior hidden, boxrule=0pt, frame hidden, enhanced, breakable, sharp corners]}{\end{tcolorbox}}\fi

@@

\caption{\label{fig-wkhighD}How the 2D model lift into high dimensions?
(a) visualize the points and the hexagonal bin centroids related 31th
and 239th hexagons, (b) visualization of the edge connected the 31th and
239th hexagons (colored in red) in the triangular mesh. A video of tour
animation is available at (need to add the link of lg\_high).}

@@

{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-diagnosticpltScurve-1.pdf}

@@

\caption{\label{fig-diagnosticpltScurve}Goodness of fit statistics from
different NLDR techniques applied to training S-curve dataset. What is
the best NLDR technique to represent the original data in 2D?}

@@

\caption{\label{fig-diagnosticpltDiffParam}Goodness of fit statistics
from different n\_neighbors parameter of UMAP applied to training
S-curve dataset. What is the best parameter choice in UMAP to represent
the original data in 2D?}

@@

{\centering \includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-diagnosticpltGau-1.pdf}

\caption{\label{fig-diagnosticpltGau}Goodness of fit statistics from
different NLDR techniques applied to training five spherical Gaussian
cluster dataset. What is the best NLDR technique to represent the
original data in 2D?}

@@

Next, we employed the UMAP technique with default parameter settings. As
illustrated in Figure~\ref{fig-pbmc}, the cell types B and Platelet are
well-separated in the UMAP layout. Moreover, CD14+ Mono, FCGR3A+ Mono,
and DC form a distinct cluster, while Naive CD4 T, NK, Memory CD4 T, and
CD8 T are grouped together in another cluster. The values utilized to
construct the smooth low-dimensional manifold are presented in
\textbf{?@tbl-table02}. The linked video, demonstrating the tour with
the model, showcases the generation of a smooth surface for this
application, enabling a comprehensive exploration of the data's
structure and relationships (see \textbf{?@fig-pbmc\_sc}).

@@

@@

\caption{\label{fig-diagnosticpltPBMC}Goodness of fit statistics from
different NLDR techniques applied to training PBMC dataset. What is the
best NLDR technique to represent the original data in 2D?}

@@

\caption{\label{fig-diagnosticpltZEI}Goodness of fit statistics from
different NLDR techniques applied to training Zeisel mouse brain
dataset. What is the best NLDR technique to represent the original data
in 2D?}
